---
title: "Chapter 4: What is This Stuff Called Probability"
output: html_document
---

### Exercise 4.1
```{r}
show(HairEyeColor) # Show data
EyeHairFreq <- apply(HairEyeColor, c("Eye", "Hair"), sum) # Sum across sex
EyeHairProp <- EyeHairFreq / sum(EyeHairFreq) # joint proportions, Table 4.1
show(round(EyeHairProp, 2))
```
First sum all attributes (`Eye`, `Hair`) along the sex dimesion. We are left with
eye and hair frequencies irrespective of `Sex`. Then we divide by the total
count get relative proportions. Lastly we print the data to decimal places.
```{r}
HairFreq <- apply(HairEyeColor, c("Hair"), sum) # Sum across sex and eye
HairProp <- HairFreq / sum(HairFreq) # marginal proportions, Table 4.1
show(round(HairProp, 2))
```
Sum all attributes (`Eye`, `Sex`) leaving the `Hair` counts, this is the marginal
of `Hair` color. Then we divide by the total count get relative proportions.
Lastly we print the data to decimal places.
```{r}

EyeFreq <- apply(HairEyeColor, c("Eye"), sum) # Sum across sex and eye
EyeProp <- EyeFreq / sum(EyeFreq) # marginal proportions, Table 4.1
show(round(EyeProp, 2))
```
Sum all attributes (`Hair`, `Sex`) leaving the `Eye` counts, this is the marginal
of `Eye` color. Then we divide by the total count get relative proportions.
Lastly we print the data to decimal places.
```{r}
EyeHairProp["Blue", ] / EyeProp["Blue"] # conditional prob, Table 4.2
```
The probability of getting blue eyes given hair color. Computed by extracting
the row of blue eyes for given hair color and diving it by the marginal value
of blue eyes.

### Exercise 4.2
```{r}
set.seed(0)
N = 500
pHeads <- 0.8
flipSequence <- sample(x=c(0,1), prob=c(1-pHeads,pHeads), size=N, replace=TRUE)
r <- cumsum(flipSequence)
n <- 1:N
runProp <- r/n
plot(n,runProp, type="o", log="x", col="skyblue",
     xlim=c(1,N), ylim=c(0.0,1.0), cex.axis=1.5,
     xlab="Flip Number", ylab="Proportion Heads", cex.lab=1.5,
     main="Running Proportion of Heads", cex.main=1.5)
abline(h=pHeads, lty="dotted")
flipLetters <- paste(c("T", "H")[flipSequence[1:10]+1], collapse="")
displayString <- paste0("Flip Sequence = ", flipLetters, "...")
text(N, .9, displayString, adj=c(1,0.5), cex=1.3)
text(N, .8, paste("End Proportion =", runProp[N]), adj=c(1,0.5), cex=1.3)
```

### Exercise 4.3
```{r}
4*2/48.
2*4*2/48.
```

### Exercise 4.4
**(A)**
```{r}
xlo <- 0.0
xhigh <- 1.0
dx <- 0.01
x <- seq(from=xlo, to=xhigh, by=dx)
y <- 6*x*(1-x)
plot(x,y, type="h", lwd=1, cex.axis=1.5,
     xlab="x", ylab="p(x)", cex.lab=1.5,
     main="Normal probability Density",
     cex.main=1.5)
lines(x, y, lwd=3, col="skyblue")
area <- sum(y*dx)
text(0.8, 0.9*max(y), bquote(paste(Delta, "x= ", .(dx))),
     adj=c(0,.5), cex=1.2)
text( 0.8 , .75*max(y) ,
      bquote(
        paste( sum(,x,) , " " , Delta , "x p(x) = " , .(signif(area,3)) )
      ) , adj=c(0,.5) , cex=1.2 )
```

**(B)**

$\int_0^{1}p(x)dx=\int_0^{1} 6x(1-x)dx = (3x^2 - 2x^3) \Big|_0^1$ = 1

**(C)**
Yes the equation satisfies the normalized equation by evaluating to 1.0.

**(D)**
From the graph the max value appears to be 1.5 which can be verified by
$p(0.5)=6(0.5)(1-0.5)=1.5$

### Exercise 4.5
**(A)**
```{r}
meanval = 0.0               
sdval = 0.2                 
xlow  = meanval - sdval 
xhigh = meanval + sdval 
dx = sdval/100.               
x = seq( from = xlow , to = xhigh , by = dx )
y = ( 1/(sdval*sqrt(2*pi)) ) * exp( -.5 * ((x-meanval)/sdval)^2 )
plot( x , y , type="h" , lwd=1 , cex.axis=1.5
        , xlab="x" , ylab="p(x)" , cex.lab=1.5 ,
        , main="Normal Probability Density" , cex.main=1.5 )
lines( x , y , lwd=3 ,  col="skyblue" )
area = sum( dx * y )
text( .09 , max(y) , bquote( paste(Delta , "x = " ,.(dx)) )
      , adj=c(0,.5) , cex=1.2 )
text( .09 , .95*max(y) ,
      bquote(
        paste( sum(,x,) , " " , Delta , "x p(x) = " , .(signif(area,3)) )
      ) , adj=c(0,.5) , cex=1.2 )
```

**(B)**
Form the data the mean has value $\mu = 162$ cm and using $\mu + \sigma = 177$,
$\sigma$ has the value $\sigma = 15$ cm.

### Exercise 4.6
**(A)**
```{r}
ice.cream <- c(0.3, 0.6, 0.3)
fruit <- c(0.6, 0.3, 0.1)
french.fries <- c(0.1, 0.1, 0.6)
data <- data.frame(ice.cream, fruit, french.fries)
p.grade <- c(0.2, 0.2 , 0.6)
data * p.grade
```